Spurious Sentience:\\ Dovelating Inflated AI World Model Claims \raisebox{-6pt}{\includegraphics[scale=0.09]{dove_1f54a-fe0f.png}}

### 2024-01-26 main.tex
Updates from Overleaf

Recently, Google DeepMind made claims their AlphaGeometry model reached a 'milestone' towards AGI. This model has the ability to solve complex geometry problems, which, they state it does without the need for human demonstrations during training. Although this is a fantastic achievement by the team, where we observe the model outperforming previous state of the art (SOTA) models, the underlying model is far from AGI. As discussed in the paper \cite{}, AlphaGeometry utilizes a rule-bound deduction engine, allowing the use of an NLP model for further suggestions when it is unable to locate a logical answer. With this combination, it was able to score highly (\textit{25/30} questions solved) where the gold standard human equivalent was \textit{25.9/30}.

However, it is worth noting, that work such as this had been initially introduced in the 1950s. Without the use of an LLM, logical inference systems proved 100\% accurate in proving all the theorems of Euclidean Geometry. They were able to achieve this success due to geometry being an axiomatically closed system. Therefore, despite DeepMinds success in creating a powerfully fast Geometry solving machine, it is still far from being an AGI.

One of the more contentious pieces of literature recently studied is the paper by \cite{gurnee2023language}

@@

@@

\subsection{Example: Deep Learning}\label{example-deep-learning}

@@

\includegraphics[width=0.5\textwidth]{results/figures/dl.png}

@@

\caption{\label{fig-dl-results}The top panel shows the actual GDP growth
and fitted values from the autoencoder model. The bottom panel shows the
observed average level and spread of the yield curve (solid) along with
the predicted values from the linear probe based on the latent embedding
(dashed).}

@@

Instead of computing a measure based on predicted labels, we can use linear probes to assess if the fine-tuned model has learned associative patterns between central bank communications and key economic indicators. To this end, we have further pre-processed the data provided by \cite{shah2023trillion} and used their proposed model to compute layer-wise embeddings for all available sentences. We have made these available and easily accessible through a small Julia package: \href{https://github.com/pat-alt/TrillionDollarWords.jl}{TrillionDollarWords.jl}. For each layer, we have then computed linear probes on two inflation indicators---the Consumer Price Index (CPI) and the Producer Price Index (PPI)---as well as US Treasury yields at different levels of maturity.

>>>>

Recently, Google DeepMind made claims their AlphaGeometry model \cite{trinh2024geometry} reached a 'milestone' towards AGI. This model has the ability to solve complex geometry problems, which, they state it does without the need for human demonstrations during training. Although this is a fantastic achievement by the team, where we observe the model outperforming previous state of the art (SOTA) models, the underlying model is far from AGI. As discussed in the paper \cite{trinh2024geometry}, AlphaGeometry utilizes a rule-bound deduction engine, allowing the use of an NLP model for further suggestions when it is unable to locate a logical answer. With this combination, it was able to score highly (\textit{25/30} questions solved) where the gold standard human equivalent was \textit{25.9/30}.

However, as noted by Hector Zenil \cite{hector2024linkedin}, that work such as this had been initially introduced in the 1950s. Without the use of an LLM, logical inference systems proved 100\% accurate in proving all the theorems of Euclidean Geometry. They were able to achieve this success due to geometry being an axiomatically closed system. Therefore, despite DeepMinds success in creating a powerfully fast Geometry solving machine, it is still far from being an AGI.

One of the more contentious pieces of literature recently studied is the paper by \cite{gurnee2023language}. This paper, which was the inspiration for our work, received numerous points of critical feedback when they discussed their work on the X\footnote{https://twitter.com} (formally Twitter) platform. Although not directly mentioning AGI, they do hint at the emergence of intelligence, specifically with the creation of the world model.

They claim said models generate an internal representation of the world, with which spatiotemporal representations are created. Further to this we see claim to the generalization of tasks, in which the model can answer questions regarding spatial and temporal relationships, without the need for further training. Although these can be seen as building blocks for a more comprehensive model, it is also fair to assume that it is mere conjecture and more likely spurious correlations.

@@

\subsection{Example: Injecting a World Model}\label{example-deep-learning}

\includegraphics[width=0.5\textwidth]{results/figures/map.png}

\caption{\label{fig-map}Mapping MNIST digits to FIFA World Rankings.}

@@

\subsection{Example: Autoencoder}\label{example-deep-learning}

@@

\includegraphics[width=1.0\textwidth]{results/figures/dl.png}

@@

\caption{\label{fig-dl-results}The left chart shows the actual GDP growth
and fitted values from the autoencoder model. The right chart shows the
observed average level and spread of the yield curve (solid) along with
the predicted values (in-sample) from the linear probe based on the latent embeddings
(dashed).}

@@

Instead of computing a measure based on predicted labels, we can use linear probes to assess if the fine-tuned model has learned associative patterns between central bank communications and key economic indicators. To this end, we have further pre-processed the data provided by \cite{shah2023trillion} and used their proposed model to compute layer-wise embeddings for all available sentences. We have made these available and easily accessible through a small Julia package: \href{https://anonymous.4open.science/r/TrillionDollarWords/README.md}{TrillionDollarWords.jl}. For each layer, we have then computed linear probes on two inflation indicators---the Consumer Price Index (CPI) and the Producer Price Index (PPI)---as well as US Treasury yields at different levels of maturity.
